\chapter{Algorithmic Approach I: Label Setting and Bounding Foundations}
\label{chap:label_setting_approach}

\section{The Time-Dependent Transportation Graph}
\label{sec:time_dependent_graph}

With a fully normalized, UTC-synchronized dataset of European transit schedules, the next step was constructing the mathematical environment for the routing algorithm. A standard static spatial graph $G = (V, E)$ is fundamentally inadequate for modeling scheduled public transportation, as an edge implies continuous availability.

\medskip

To rigorously test different modeling paradigms, the first architectural approach evaluates the network as a time-dependent multi-graph. In this AoE architecture, the vertices remain purely spatial, while the dimension of time is encoded directly into the attributes of the edges. This approach was deliberately chosen to maintain a highly compact memory footprint, avoiding the node explosion inherent to fully time-expanded models.

\medskip

The graph was instantiated using the NetworkX library (see library documentation in \cite{hagberg2008exploring}) with the following structural rules:

\begin{itemize}
    \item \textbf{Spatial Nodes ($V$):} A single vertex was created for each unique physical location. Nodes were tagged with specific attributes: their sovereign country (critical for the objective function) and their infrastructure type (airport or station). 
    \item \textbf{Temporal Transit Edges ($E$):} Flights and trains were modeled as directed edges between origin and destination nodes. Because multiple trains or flights travel between the same two cities throughout the day, these edges stored dynamic temporal attributes. Instead of a single static weight, each edge contained an array of valid (\texttt{departure\_time}, \texttt{arrival\_time}) pairs, formatted in absolute UTC minutes. To account for the 48-hour Guinness World Record window, train schedules (which operate on daily cycles) were programmatically duplicated, adding $1440$ minutes to populate the second day of the graph.
    \item \textbf{Static Transfer Edges:} The manually validated intra-city connections (e.g., traversing from a city's airport to its central rail station) were integrated as standard, bidirectional static edges. Unlike transit edges, transfer edges were assigned a fixed scalar weight representing the human connection time in minutes, as they can be traversed at any point during the day.
\end{itemize}

By compressing schedules into arrays on spatial edges, this AoE architecture successfully avoided node redundancy. However, this time-dependent structure shifts the chronological burden entirely onto the search algorithm, which must dynamically iterate through edge arrays to find valid connections based on its arrival time at any given node. This compact graph serves as the baseline environment to test the exact optimization methods detailed in the following sections.

\section{Algorithm Evolution: From Baseline to Bounding}
\label{sec:algorithm_evolution}

\noindent
Because the Guinness World Record challenge is fundamentally a multi-objective optimization problem (maximizing unique countries visited while minimizing time within a strict 24-hour limit) a standard shortest-path algorithm like Dijkstra's cannot be used. Instead, the search must be modeled using DP via a Label-Setting algorithm (see Algorithm 1 in Appendix B).

\medskip

In this paradigm, as the algorithm traverses the time-dependent multi-graph, it assigns a state ``label'' $L$ to each node. A label is defined by the tuple $L = (v, t, C, s)$. Here, $v$ tracks the current node, $t$ is the current absolute minute, $C$ represents the mathematical set of unique countries visited so far, and $s$ denotes \texttt{start\_time}, which marks when the journey began. The cardinality of the set, denoted as $|C|$, represents the integer count of unique countries visited.

\medskip

To prevent the search space from expanding infinitely, the algorithm relies on Pareto Dominance. At any given node, a new label $L'$ is strictly dominated by an existing label $L''$ if the existing path arrived earlier or at the same time ($L''.t \leq L'.t$) while having visited a superset or equal set of countries ($L''.C \supseteq L'.C$), with at least one condition being strictly better. Dominated labels are instantly discarded.

\medskip

However, implementing this baseline algorithm revealed a fundamental limitation: the Pareto dominance conditions were too relaxed. Across a network spanning roughly 40 countries, the combinatorial permutations of visited country sets $C$ caused an explosion of non-dominated labels, rendering the baseline approach computationally unfeasible. This prompted a series of algorithmic refinements, where mathematical pruning techniques were progressively introduced to control the state space.

\subsection{Reachability and Upper Bound (UB) Pruning}
\label{subsec:baseline_label_setting}

The first major architectural change was introducing predictive pruning (see Algorithm 2 in Appendix B). Before expanding a label, the algorithm calculated an optimistic Upper Bound ($UB$) of the maximum new countries that could theoretically be reached from the current node $v$ within the remaining time.

\medskip

To make this computationally tractable without running a full sub-search at every node, time was discretized into fixed intervals, or ``buckets'' $b$. For any given node $v$ and time bucket $b$, a precomputation step (detailed in Algorithm 3, Appendix B) generated $R(v, b)$, the absolute set of all sovereign countries reachable from that location before the 24-hour clock expired. The theoretical Upper Bound was then calculated as the cardinality of the set difference between the reachable countries and the already visited countries: $UB = |R(v, b) \setminus C|$.

\medskip

If the current number of countries visited plus this theoretical maximum ($|C| + UB$) was strictly less than or equal to the current global record, or couldn't beat the existing record, the path was mathematically dead and immediately pruned.

\subsection{Time-Budgeted UB and Minimum Transit Time (MTT)}
\label{subsec:ub_pruning}

While standard $UB$ pruning successfully removed hopeless paths, it was overly optimistic. It assumed that if 10 new countries were reachable from a node, all 10 could be visited, completely ignoring the physical time required to travel between them.

\medskip

To tighten this bound, the concept of MTT was introduced. The $MTT$ represents the absolute shortest physical time required to enter, traverse, and exit a specific country's transit network. For instance, even under perfect schedule conditions, traveling across Switzerland or Germany requires a mathematically provable minimum number of minutes. The algorithm precomputed this $MTT$ for every country in the dataset prior to the main search execution (see Algorithm 5 in Appendix B).

\medskip

When calculating the $UB$, the algorithm sorted all unvisited, reachable countries from the fastest $MTT$ to the slowest (see Algorithm 6 in Appendix B). It then sequentially added these minimum transit times to a \texttt{time\_spent} counter. The moment the accumulated \texttt{time\_spent} exceeded the actual hours remaining in the 24-hour journey, the algorithm stopped incrementing the $UB$. This ``Time-Budgeted UB'' improved convergence speed: it accurately pruned paths that possessed theoretical geographic connectivity but lacked the physical time budget required to actually execute the transits.

\subsection{Priority Scoring, Hard Floors, and Rollouts}
\label{subsec:priority_scoring_hard_floors_rollouts}

To further optimize the search, the standard queue was replaced with a Priority Queue. Rather than expanding blindly, labels were ranked using a custom scoring formula. The formula was designed to prioritize deep, fast routes. For example, a core heuristic evaluated ${Score}(L) = |C| \times 10000 - t$, effectively forcing the algorithm to expand labels with the highest country count, while using an earlier absolute time $t$ as the mathematical tie-breaker to favor faster connections (see Algorithm 4 in Appendix B).

\medskip

Despite prioritized expansion, execution logs revealed the algorithm still wasted resources expanding labels near the end of the timeline that had only accumulated a few countries. To counter this, a Global Hard Floor was introduced (see Algorithm 9 in Appendix B). By establishing a GlobalMinTime (the minimum realistic time to visit any single country), the algorithm calculated the physical time required to bridge the gap between $|C|$ and the target record. If this required time exceeded the remaining clock, the label was killed. Crucially, all arbitrary parameters for these heuristics (such as setting the GlobalMinTime to just 25 minutes) were deliberately defined to be overly optimistic. The architectural philosophy was to tolerate redundant computational overhead rather than risk pruning a mathematically viable, record-breaking route.

\medskip

To capitalize on the Priority Queue, a Monte-Carlo Rollout mechanism was injected (see Algorithm 8 in Appendix B). Every 1,000 iterations, the algorithm paused standard expansion and launched a rapid, randomized ``smart-greedy'' probe deep into the graph (see Algorithm 7 in Appendix B). Using heavily biased weights to favor unvisited countries and a Tabu list to prevent local loops, these rollouts operated in milliseconds. If a rollout successfully reached the end of the day and beat the current global record, the new record was dynamically adopted, allowing the updated bound to prune a large portion of the priority queue immediately.

\subsection{The Bi-directional Search Pivot}
\label{subsec:bidirectional_search_pivot}

The final algorithmic evolution within this paradigm was motivated by the failure modes of the Monte-Carlo rollouts. Debugging revealed that rollouts were highly ineffective when they reached the night hours; physical train schedules became incredibly sparse, trapping the greedy probes in geographic dead-ends.

\medskip

To bypass this nocturnal sparsity, the search was split into two halves. The architecture was rewritten as a Bi-directional Search: one algorithm ran forward in time from the start of the day, while an identical algorithm ran backward in time from the end of the day (see Algorithms 10 and 11 in Appendix B). Both halves executed for 14 hours, creating a 4-hour overlap in the middle of the day.

\medskip

Finally, a dedicated ``Stitcher'' algorithm evaluated the intersection of these two sets (see Algorithm 12 in Appendix B). By iterating through overlapping nodes and validating time continuity and geographic unions ($C_{\textsl{forward}} \cup C_{\textsl{backward}}$), the stitcher could reconstruct a globally optimal 24-hour path. This parallel, bi-directional approach effectively bypassed the sparse night-time bottlenecks and maximized graph coverage during the dense daytime schedules.

\section{The Dimensionality Collapse}
\label{sec:dimensionality_collapse}

Despite these successive refinements, from baseline Pareto dominance to time-budgeted Upper Bounds, hard floors, and bi-directional stitching, the architecture ultimately could not handle the combinatorial complexity. It is relevant to clarify that the algorithm did not fail due to an Out-Of-Memory (OOM) error. Rather, it hit a computational bottleneck.

\medskip

The root of this bottleneck lay in the maintenance of the Pareto frontiers. As the search progressed deeper into the time-dependent multi-graph, the frontiers at each node grew increasingly dense. Because the Guinness World Record routing is fundamentally multi-objective (minimizing time versus maximizing unique countries), vast numbers of paths are mathematically incomparable. A route that visits 8 countries by 14:00 cannot dominate a route that visits 9 countries by 16:00. Consequently, both labels must be kept alive. As these non-dominated sets grew exponentially, the algorithm was forced to perform extensive comparative loops for every newly generated label just to establish dominance, making queue operations prohibitively expensive.

\medskip

This inefficiency was heavily compounded by the execution environment. The algorithm was executed locally on a single machine rather than being deployed to an HPC cluster. This was not a limitation of resources, but a deliberate architectural necessity. Dynamic Programming and Label-Setting algorithms fundamentally rely on continually evaluating and updating a globally shared state, specifically, the dynamic Pareto frontiers at every node in the graph. Attempting to parallelize this specific architecture across distributed cluster nodes (which operate on isolated memory pools) would have introduced massive IPC overhead. The constant network latency required to safely lock and synchronize the frontiers across dozens of different parallel workers would likely have negated the benefits of parallelization. Thus, the approach was inherently bottlenecked by single-thread CPU execution speeds.

\medskip

After weeks of continuous local execution, the algorithm struggled to converge on any record-breaking paths, a limitation that the AoN approach would later address. It is worth acknowledging that performance could theoretically have been pushed further. By implementing stricter dominance rules, more non-optimistic heuristics (such as aggressively killing technically viable paths to artificially force the queue to empty faster), or artificially forcing the queue to empty faster, the algorithm could have potentially achieved faster convergence. However, doing so would have completely sacrificed the exactness of the search. Given the highly specific, non-intuitive nature of train schedules, overly aggressive pruning risked the blind deletion of the optimal, record-winning route.

\medskip

More importantly, the ultimate objective of this research was not to infinitely tune a single algorithm, but to conduct a comparative evaluation between two fundamental modeling paradigms. After completing several major evolutionary iterations of the algorithm running on the AoE architecture, its core computational trade-offs were conclusively established: it is highly memory-efficient, but it shifts an excessive processing burden onto the search algorithm. The analysis now transitions to evaluating the second architectural candidate: the highly parallelizable AoN framework, explored in the following chapter.
